---
title: "Lab 3"
author: "Alex Brown"
date: "1/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)
```

## Binary logistic regression: Penguins!

Read in the data
```{r}
gentoo <- read_csv("gentoo_lter.csv")
chinstrap <- read_csv("chinstrap_lter.csv")

penguins <- full_join(chinstrap, gentoo) %>% 
  clean_names() %>% 
  mutate(sex = str_to_lower(sex)) %>% 
  filter(sex %in% c("male", "female"))
```

Now let's graph this:

```{r}
ggplot(data = penguins, aes(x = sex, y = body_mass_g)) +
  geom_jitter(
    aes(color = sex,
        pch = sex)
  ) +
  facet_wrap(~species)
```

Code Chinstrap penguins as (1) and Gentoo penguins as (0):

```{r}
penguin_bin <- penguins %>% 
  mutate(
    sp_binary = case_when(
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ 1,
      species == "Gentoo penguin (Pygoscelis papua)" ~ 0
    )
  )
```

Now let's run our binary logistic regression:

```{r}
penguin_blr <- glm(sp_binary ~ sex + body_mass_g,
                   family = binomial,
                   data = penguin_bin)

summary(penguin_blr)
```

Let's make some predictions for penguins (classification...kinda):

1. What is the probability that a penguin is a Chinstrap penguin, if it weighs 4500 grams and is male? 
```{r}
df_m4500 <- data.frame(sex = "male", body_mass_g = 4500)

# Find the log odds of this penguin being Chinstrap: 
m4500_logodds <- predict(penguin_blr, newdata = df_m4500, type = "link")
m4500_logodds

m4500_prob <- predict(penguin_blr, newdata = df_m4500, type = "response")
m4500_prob
```
2. What is the probability that a penguin is Chinstrap, if it is recorded as female of 4000 g?

```{r}
df_f4000 <- data.frame(sex = "female", body_mass_g = 4000)

f4000_prob <- predict(penguin_blr, newdata = df_f4000, type = "response")
f4000_prob
```

Now let's make an entire data frame and use it to make then visualize logistic regression outcomes:

```{r}
penguins_mock <- data.frame(
  body_mass_g = rep(seq(3000, 6000, length = 200), 2),
  sex = c(rep("male", 200), rep("female", 200))
)
```

Now let's use those to make predictions for all of them:

```{r}
full_predict <- predict(penguin_blr, newdata = penguins_mock, type = "response", se.fit = TRUE)

# Coerce predictions into a data frame, with the mock data + SEs:
final_df <- data.frame(penguins_mock, 
                       full_predict$fit, 
                       full_predict$se.fit)

colnames(final_df) <- c("penguin_mass", "sex", "probability", "se")
```

Now let's graph it!
```{r}
ggplot(data = final_df, aes(x = penguin_mass, y = probability)) +
  geom_line(aes(color = sex)) +
    scale_x_continuous(limits = c(3500, 5500)) +
  geom_ribbon(aes(ymin = probability - se, ymax = probability + se, fill = sex), alpha = 0.3) +
  labs(x = "Penguin body mass (g)",
       y = "Probability of being Chinstrap") +
  theme_minimal()
```

Some questions to think about: 

- What are some assumptions we've made? 
- What are the limitations? 
- Do these align with what we'd expect from our data visualization?

# PART 2. (Re)introduction to visualizing spatial data

The `cougar_connections` folder contains spatial data for mountain lion connectivity in the Sierra Nevada foothills. From the researchers: "The least-cost corridors identify the best swath of habitat available for focal species to move from one landscape block to another based on predicted suitable habitat as identified by the northern Sierra Nevada foothills wildlife connectivity project."

**Citation:** Beier, P., D. R. Majka, and J. S. Jenness. 2007. Conceptual steps for designing wildlife corridors.

Link to project report: https://nrm.dfg.ca.gov/FileHandler.ashx?DocumentID=85358

### Step 1. Read in the data!

Use `sf::read_sf()` to read in the spatial data together!
```{r}
cougars <- read_sf(dsn = here("cougar_connections"), layer = "ds1014") %>%
  clean_names()
```

Notice that the last column is "geometries" - that contains the *sticky* spatial data - meaning that we can wrangle & work with our data just like we've done before, but the spatial information will stick to it! 

### STEP 2: Practice some spatial wrangling with sticky geometries

For example: Let's say we're only interested in really large corridors with areas greater than 24,000 acres. We can filter the dataset normally, and only keep that attribute...but the geometry will stick to it! 

```{r}
# Use wrangling functions as usual (+ pipe, etc!)
large_corridors <- cougars %>% 
  filter(area_ac > 20000) %>% 
  select(area_ac)

# Look at 'large_corridors' data frame and see that the geometries column still appears, even though we didn't explicitly select it. Cool sf magic! 

# We can even use the base plot() function to see what's happening (careful of doing this if we have multiple attributes):
plot(large_corridors)
```

Let's create a different subset from the original data that contains the following variables:

- elev_mean (mean elevation)
- ecoregs (eco-regions)
- pc_urban (percent urban area in polygon)

```{r}
cougar_sub <- cougars %>% 
  select(elev_mean, ecoregs, pc_urban)
```

### STEP 3. Visualizing spatial polygons with `geom_sf`
```{r}

ggplot(data = cougar_sub) +
  geom_sf(aes(fill = elev_mean), color = NA)

```

### STEP 4. Actually give it some context & interactivity! 

We can use `tmap` to make a static or interactive plot over an existing basemap.
```{r}
# This will make a stationary (non-interactive) plot
tmap_mode("plot") # switch to "view" to make interactive! 

# 
tm_shape(cougar_sub) +
  tm_fill("elev_mean")
```

OK maybe not the most inspiring thing, but it works...now let's change the view mode: 
```{r}
tmap_mode("view") # Set to interactive viewing
tm_shape(cougar_sub) +
  tm_fill("elev_mean") +
  tm_basemap("Stamen.Terrain")
```

See http://leaflet-extras.github.io/leaflet-providers/preview for a preview of basemaps you can add with `tm_basemap`!

Try out some different basemaps and test interactivity, and see that it's maintained when you knit! 
